---
title: "Part 2"
output: html_notebook
---

# Setup Notebook

```{r setup, results='hide'}
knitr::opts_knit$set(root.dir = '..')
devtools::install_github("tidyverse/ggplot2") #version with stat_qq_line

# Load requirements
library(dplyr)
library(tidyr)
library(ggplot2)
```

# How much data is available?

The dataset is a wrangled version of the 2010 common core dataset from NCES. The wrangling steps are documented in [Part_1.ipynb](https://github.com/albemlee/schooldistrictclassification/blob/master/notebooks/Part_1.ipynb). There should be 365 variables and 18,175 observations.

```{r, results='show'}
# Load data
data <- read.csv('outputs/encoded_data_2010.csv')
data <- data %>% mutate_if(is.integer,as.numeric)
print(dim(data))
```

# How much data is missing?

Prior to this analysis, the dataset was already wrangled and variables with more than half the values missing were removed. As a sanity check, I made sure the wrangled dataset does not contain too many missing values.

```{r}
print(paste('Total Cells', 18175*177, sep=': '))
print(paste('Total Missing Cells', sum(is.na(data)), sep=': '))
print(paste('Proportion of Cells Missing', sum(is.na(data))/(18175*356), sep=': '))
```

About 5 percent of the cells are missing. Next, I want to know which columns have the most missing values.

```{r}
missing_value_counts = c()

for(column in names(data)){
  missing_value_counts <- c(missing_value_counts, (sum(is.na(select(data, column)))))
}

missing_value_counts = data.frame(names(data), missing_value_counts)
names(missing_value_counts) <- c('column', 'missing_value_count')
print(
    missing_value_counts %>%
      select(column, missing_value_count) %>%
      arrange(desc(missing_value_count)) %>%
      top_n(5)
)
    
```

Let's see the distribution of missing values.

```{r}
ggplot(missing_value_counts, aes(missing_value_count)) +
  geom_histogram(binwidth=500)
```

```{r}
missing_value_counts %>%
  filter(missing_value_count < 2000) %>%
  select(column)
```

I will be focusing on features with less than 2,000 missing values.

```{r}
clustering_set <- data %>% select(
  total_schools,
  administrators_district,
  administrators_school,
  teachers_total,
  teachers_elementary,
  teachers_secondary,
  staff_student_support,
  special_education_students,
  english_language_learners,
  white_students,
  white_male_students,
  white_female_students,
  hispanic_students,
  black_non_hispanic_students,
  librarian_media_specialists,
  hispanic_male_students,
  hispanic_female_students,
  black_non_hispanic_male_students,
  instructional_aides,
  black_non_hispanic_female_students,
  total_students,
  total_revenue,
  total_federal_revenue,
  federal_revenue_state_title_i,
  federal_revenue_ideas,
  federal_revenue_math_science_quality,
  federal_revenue_drug_free,
  federal_revenue_vocational_tech_training,
  federal_revenue_thru_state_bilingual_education,
  federal_revenue_thru_state_other,
  federal_revenue_thru_state_child_nutrition_program,
  federal_revenue_nonspecified,
  federal_revenue_impact_aid,
  federal_revenue_indian_education,
  federal_revenue_other,
  total_state_revenue,
  state_revenue_general_formula_assistance,
  state_revenue_staff_improvement,
  state_revenue_special_education,
  state_revenue_compensatory_basic_training,
  state_revenue_bilingual_education,
  state_revenue_gifted_talented,
  state_revenue_vocational_programs,
  state_revenue_school_lunch,
  state_revenue_capital_outlay_debt_services,
  state_revenue_transportation,
  state_revenue_other,
  state_revenue_nonspecified,
  state_revenue_employee_benefits,
  state_revenue_not_employee_benefits,
  total_local_revenue,
  local_revenue_other_school_systems,
  local_revenue_cities_counties,
  local_revenue_tuition_fee_pupils_parents,
  local_revenue_transportation_fee_pupil_parents,
  local_revenue_school_lunch,
  local_revenue_textbook_sale_rental,
  local_revenue_district_activity_receipt,
  local_revenue_student_fee_nonspecified,
  local_revenue_other_sales_services,
  local_revenue_rent_royalties,
  local_revenue_property_sales,
  local_revenue_interest_earnings,
  local_revenue_fines_forfeits,
  local_revenue_private_contributions,
  local_revenue_miscellaneous,
  local_revenue_NCES,
  total_expenditure,
  total_expenditure_elementary_secondary,
  total_expenditure_instruction,
  total_expenditure_instruction_public,
  expenditure_private_school,
  expenditure_charter_school,
  total_expenditure_support_services,
  expenditure_support_services_pupils,
  expenditure_support_services_instructional_staff,
  expenditure_support_services_general_administration,
  expenditure_support_services_school_administration,
  expenditure_support_services_maintenance,
  expenditure_support_services_transportation,
  expenditure_support_services_business,
  expenditure_support_services_nonspecified,
  total_expenditure_other_elementary_secondary,
  expenditure_food_service,
  expenditure_enterprise,
  expenditure_other_elementary_secondary,
  total_expenditure_non_elementary_secondary,
  expenditure_non_elementary_secondary_community_service,
  expenditure_non_elementary_secondary_adult_education,
  expenditure_non_elementary_secondary_other,
  total_expenditure_capital_outlay,
  expenditure_capital_outlay_construction,
  expenditure_capital_outlay_land_existing_structures,
  expenditure_capital_outlay_instructional_equipment,
  expenditure_capital_outlay_other_equipment,
  expenditure_capital_outlay_nonspecified,
  payments_state_government,
  payments_local_government,
  payments_other_school_systems,
  interest_on_debt,
  total_salaries,
  salaries_instruction,
  salaries_regular_education,
  salaries_special_education,
  salaries_vocationall_education,
  salaries_other_education,
  salaries_support_services_pupils,
  salaries_support_services_instructional_staff,
  salaries_support_services_general_administration,
  salaries_support_services_school_administration,
  salaries_support_services_maintenance,
  salaries_support_transportation,
  salaries_support_services_business,
  salaries_food_service,
  total_employee_benefits,
  employee_benefits_instruction,
  employee_benefits_support_services_pupil,
  employee_benefits_support_services_instructional_staff,
  employee_benefits_support_services_general_administration,
  employee_benefits_support_services_school_administration,
  employee_benefits_support_services_maintenance,
  employee_benefits_support_transportation,
  employee_benefits_support_services_business,
  employee_benefits_food_service,
  employee_benefits_enterprise,
  textbooks,
  long_term_debt_outstanding_beginning_fiscal_year,
  long_term_debt_issued_during_fiscal_year,
  long_term_debt_retired_during_fiscal_year,
  long_term_debt_outstanding_end_fiscal_year,
  short_term_debt_outstanding_beginning_fiscal_year,
  short_term_debt_outstanding_end_fiscal_year,
  ARRA_revenue,
  ARRA_current_expenditures,
  ARRA_capital_outlay_expenditures,
  asian_hawaiian_native_pacific_islander_students,
  american_indian_alaskan_native_students,
  asian_hawaiian_native_pacific_islander_male_students,
  asian_hawaiian_native_pacific_islander_female_students,
  american_indian_alaskan_native_male_students,
  exist_five_years
)

clustering_set <- na.omit(clustering_set)

for (i in 2:10){
  clust <- kmeans(select(clustering_set, -one_of("exist_five_years")), centers=i, iter.max=30)
  print(paste(i, ':', sum(clust$withinss)/clust$totss))
}

```

```{r}
clustering_set %>%
  group_by(cluster) %>%
  summarize(sum(exist_five_years), n())


```







